---
title: "Caret Package"
author: "Carlos Sanchez"
date: "05/03/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("~/R courses/8_Machine_Learning/Week2")
```
 
 
```{r}
library(kernlab)
library(caret)
library(e1071)
data(spam)
head(spam)
```


```{r}
inTrain <- createDataPartition(y=spam$type, p=0.75, list=FALSE)
training <- spam[inTrain, ]
testing <- spam[-inTrain, ]
dim(training)
```

# Caret Package

```{r}
set.seed(32343)
modelFit <- train(type ~ ., data = training, method = "glm")
modelFit$finalModel
```


```{r}
predictions <- predict(modelFit, newdata=testing)
predictions
```


```{r}
confusionMatrix(predictions, testing$type)
```


# Data Slicing

```{r}
data(spam)
inTrain <- createDataPartition(y=spam$type,
                               p=0.75, list=FALSE)
training <- spam[inTrain, ]
testing <- spam[-inTrain, ]
dim(training)
```

SPAM Example - K-Fold

#Return Training

```{r}
set.seed(123)
folds <- createFolds(y=spam$type, k=10,
                     list = TRUE, returnTrain = TRUE)
sapply(folds, length)
```


```{r}
folds[[1]][1:10]
```

#Return Test

```{r}
set.seed(123)
folds <- createFolds(y=spam$type, k=10,
                     list = TRUE, returnTrain = FALSE)
sapply(folds, length)
```


```{r}
folds[[1]][1:10]
```

#Resampling

```{r}
set.seed(123)
folds <- createResample(y=spam$type, times = 10,
                        list = TRUE)
sapply(folds, length)
```

#Time Slices

```{r}
set.seed(123)
tme <- 1:1000
folds <- createTimeSlices(y=tme, initialWindow = 20,
                          horizon = 10)
names(folds)
```


```{r}
folds$train[[1]]
folds$test[[1]]
```

#Training options

```{r}
data(spam)
inTrain <- createDataPartition(y=spam$type,
                               p=0.75, list=FALSE)
training <- spam[inTrain, ]
testing <- spam[-inTrain, ]

modelFit <- train(type ~ ., data = training, method = "glm")
```

#Plottting predictors

```{r}
library(ISLR)
library(ggplot2)
library(caret)
data(Wage)
summary(Wage)
```

```{r}
inTrain <- createDataPartition(y=Wage$wage,
                               p=0.75, list=FALSE)
training <- Wage[inTrain, ]
testing <- Wage[-inTrain, ]
dim(training)
dim(testing)
```


```{r}
featurePlot(x=training[, c("age", "education", "jobclass")],
            y= training$wage,
            plot="pairs")
```


```{r}
qplot(age, wage, data = training)
qplot(age, wage, color = jobclass, data = training)

```


```{r}
qplot(age, wage, color = education, data = training) +
  geom_smooth(method = 'lm', formula = y ~ x)

```

```{r}
library(Hmisc)
cutWage <- cut2(training$wage, g=3)
table(cutWage)
```

```{r}
p1 <- qplot(cutWage, age, data = training, fill = cutWage,
            geom = c("boxplot"))
p1
```

```{r}
library(gridExtra) 

p2 <- qplot(cutWage, age, data = training, fill = cutWage,
            geom = c("boxplot", "jitter"))
grid.arrange(p1, p2, ncol =2)
```


```{r}
t1 <- table(cutWage, training$jobclass)
t1
```


```{r}
prop.table(t1,1)
```


```{r}
qplot(wage, color=education, data=training, geom = "density ")

```

# Basic Preprocessing
```{r}
data(spam)
inTrain <- createDataPartition(y=spam$type,
                               p=0.75, list=FALSE)
training <- spam[inTrain, ]
testing <- spam[-inTrain, ]

hist(training$capitalAve, main = "", xlab = "ave. capital run length")
```


```{r}
mean(training$capitalAve)
sd(training$capitalAve)
```

```{r}
trainCapAve <- training$capitalAve
trainCapAveS <- (trainCapAve - mean(trainCapAve))/sd(trainCapAve)
mean(trainCapAveS)
sd(trainCapAveS)
```

When standardizing the test set, we will use mean and sd of training, so mean and sd will not be 0 and 1

```{r}
testCapAve <- testing$capitalAve
testCapAveS <- (testCapAve - mean(trainCapAve)) / sd(trainCapAve)
mean(testCapAveS)
sd(testCapAveS)
```


```{r}
preObj <- preProcess(training[,-58], method = c("center", "scale"))
trainCapAve <- predict(preObj, training[,-58])$capitalAve

mean(trainCapAve)
sd(trainCapAve)
```


```{r}
testCapAveS <- predict(preObj, testing[, -58])$capitalAve
mean(testCapAveS)
sd(testCapAveS)
```

Also we can standardize at the train function

```{r}
set.seed(123)
modelFit <- train(type ~ ., data = training,
                  preProcess = c("center", "scale"), method ="glm")
modelFit
```


#Box- Cox transformation

```{r}
preObj <- preProcess(training[, -58], method = c("BoxCox"))
trainCapAveS <- predict(preObj, training[, -58])$capitalAve
par(mfrow=c(1,2))
hist(trainCapAveS)
qqnorm(trainCapAveS)
```


# Imputing Data

```{r}
set.seed(123)

#Make some values NA
training$capAve <- training$capitalAve
selectNA <- rbinom(dim(training)[1], size=1, prob=0.05)==1
training$capAve[selectNA] <- NA

#Impute and standardize
preObj <- preProcess(training[, -58], method = "knnImpute")
capAve <- predict(preObj, training[, -58])$capAve

#Standardize true values
capAveTruth <- training$capitalAve
capAveTruth <- (capAveTruth-mean(capAveTruth))/sd(capAveTruth)
```



#Covariate creation



```{r}
inTrain <- createDataPartition(y=Wage$wage,
                               p=0.75, list=FALSE)
training <- Wage[inTrain, ]
testing <- Wage[-inTrain, ]
dim(training)
dim(testing)
```

```{r}
table(training$jobclass)
```

## dummy variables

```{r}
dummies <- dummyVars(wage ~ jobclass, data = training)
head(predict(dummies, newdata=training))
```

## Removing zero covariates

```{r}
nsv <- nearZeroVar(training, saveMetrics=TRUE)
nsv
```

## Transform to use cubic function

```{r}
library(splines)
bsBasis <- bs(training$age, df=3)
head(bsBasis)
```


```{r}
lm1 <- lm(wage ~ bsBasis, data = training)
plot(training$age, training$wage, pch=19, cex=0.5)
points(training$age, predict(lm1, newdata=training), col="red", pch=19, cex=0.5)
```

```{r}
head(predict(bsBasis, age=testing$age))
```


#Pre-Processing with PCA

```{r}
data(spam)
inTrain <- createDataPartition(y=spam$type,
                               p=0.75, list=FALSE)
training <- spam[inTrain, ]
testing <- spam[-inTrain, ]

M <- abs(cor(training[, -58]))
diag(M) <- 0
which(M > 0.8, arr.ind = T)
```


```{r}
plot(spam[, "num857"], spam[,"num415"])
```

```{r}
smallSpam <- spam[, c(34,32)]
prComp <- prcomp(smallSpam)
plot(prComp$x[,1], prComp$x[,2])
```


```{r}
prComp$rotation
```


```{r}
typeColor <- ((spam$type=="spam")+1 +1)
prComp <- prcomp(log10(spam[, -58] + 1))
plot(prComp$x[,1], prComp$x[,2], col=typeColor, xlab="PC1", ylab="PC2")
```

# PCA in caret

```{r}
preProc <- preProcess(log10(spam[,-58]+1), method = "pca", pcaComp = 2)
spamPC <- predict(preProc, log10(spam[,-58]+1))
plot(spamPC[,1], spamPC[,2], col = typeColor)
```


```{r}
preProc <- preProcess(log10(training[,-58]+1), method = "pca", pcaComp = 2)
trainPC <- predict(preProc, log10(training[,-58]+1))
modelFit <- train(training$type ~ ., method = "glm", data = trainPC)
```

```{r}
testPC <- predict(preProc, log10/testing[,-58]+1)
confusionMatrix(testing$type, predict(modelFit, testPC))
```


##Alternative

```{r}
modelFit <- train(type ~ ., method="glm", preProcess = "pca", data=training)
confusionMatrix(testing$type, predict(modelFit, testing))
```



# Predicting with Regression

```{r}
library(caret)
data("faithful")
set.seed(333)

inTrain <- createDataPartition(y = faithful$waiting, p=0.5, list = FALSE)
trainFaith <- faithful[inTrain, ]
testFaith <- faithful[-inTrain, ]
head(trainFaith)
```


```{r}
plot(trainFaith$waiting, trainFaith$eruptions, pch=19, col="blue", xlab="Waiting", ylab="Duration")
```



```{r}
lm1 <- lm(eruptions ~ waiting, data = trainFaith)
summary(lm1)
```


```{r}
plot(trainFaith$waiting, trainFaith$eruptions, pch=19, col="blue", xlab="Waiting", ylab="Duration")
lines(trainFaith$waiting, lm1$fitted.values, lwd=3)
```


Predict a new value

```{r}
coef(lm1)[1] + coef(lm1)[2]*80 

newdata <- data.frame(waiting=80)
predict(lm1, newdata)
```


# Predicting with Regression Multiple Covariates

```{r}
data(Wage)
Wage <- subset(Wage, select = -c(logwage))

inTrain <- createDataPartition(y=Wage$wage,
                               p=0.75, list=FALSE)
training <- Wage[inTrain, ]
testing <- Wage[-inTrain, ]
dim(training)
dim(testing)

```

```{r}
featurePlot(x=training[, c("age", "education", "jobclass")],
            y = training$wage,
            plot="pairs")
```


```{r}
qplot(age, wage, color=jobclass, data=training)
```


```{r}
qplot(age, wage, color=education, data=training)
```


```{r}
modFit <- train(wage ~ age + jobclass + education,
                method = "lm", data = training)
finMod <- modFit$finalModel
print(modFit)
```


```{r}
plot(finMod, 1, pch=19, cex=0.5, col = "#00000010")
```


```{r}
qplot(finMod$fitted, finMod$residuals, color=race, data=training)
```

Plot by index

```{r}
plot(finMod$residuals, pch=19)
```



```{r}
pred <- predict(modFit, testing)
qplot(wage, pred, color=year, data = testing)
```


```{r}
library(AppliedPredictiveModeling)
data(concrete)
library(caret)
set.seed(1000)
inTrain = createDataPartition(mixtures$CompressiveStrength, p = 3/4)[[1]]
training = mixtures[ inTrain,]
testing = mixtures[-inTrain,]
```



```{r}
library(GGally)
training$CompressiveStrength <- cut2(training$CompressiveStrength, g=3)
ggpairs(training, mapping = ggplot2::aes(colour = CompressiveStrength))
```


```{r}
library(AppliedPredictiveModeling)
data(concrete)
library(caret)
set.seed(1000)
inTrain = createDataPartition(mixtures$CompressiveStrength, p = 3/4)[[1]]
training = mixtures[ inTrain,]
testing = mixtures[-inTrain,]
```


```{r}
hist(log(training$Superplasticizer+1))
```


```{r}
library(caret)
library(AppliedPredictiveModeling)
set.seed(3433)
data(AlzheimerDisease)
adData = data.frame(diagnosis,predictors)
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]
training = adData[ inTrain,]
testing = adData[-inTrain,]
```


```{r}
train2 <- training[, grep("^IL", names(training))]
preProc <- preProcess(train2, method = "pca", thresh = 0.9)
preProc$numComp
```

```{r}
library(caret)
library(AppliedPredictiveModeling)
set.seed(3433)
data(AlzheimerDisease)
adData = data.frame(diagnosis,predictors)
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]
training = adData[ inTrain,]
testing = adData[-inTrain,]
```


```{r}
train2 <- training[, grep("^IL|diagnosis", names(training))]
preProc <- preProcess(train2, thresh = 0.8, method=c("pca"))

modelFit <- train(diagnosis ~ ., method="glm",  data=training)

```













